---
title: "LAGOS Spatial Analysis - in-class exercises"
author: "John Kemper"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
# install.packages("sf")
library(sf) #Spatial package that can read and create shapefiles 
# install.packages("mapview")
library(mapview) #Interactive maps
# install.packages("LAGOSNE")
library(LAGOSNE) #Lots and lots of clean lake data
# install.packages("USAboundaries")
library(USAboundaries) #USA states and counties
# install.packages("devtools")
library(devtools)
library(ggthemes)
library(ggplot2)

```


# LAGOS Analysis


## Loading in data

?lagos_path
?lagosne_get



### First download and then specifically grab the locus (or site lat longs)
```{r data-read}


#Lagos down
# lagosne_get(version='1.087.1',dest_folder = LAGOSNE:::lagos_path())
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

# load("lake_centers.Rdata")
lake_centers

```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states() %>%
  st_transform(2163)
View(states)

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)


mapview(minnesota)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,] %>%
    mutate(state = "Minnesota")

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r }
#Map the outlines of Iowa and Illinois
IowaIll <- states %>%
  filter(name %in% c('Iowa', 'Illinois')) %>%
  st_transform(2163)

mapview(IowaIll)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

```{r}

####Clip the LAGOS data to only include lakes that occur in the states of Iowa and Ill

IowaIll_lakes <- spatial_lakes[IowaIll,]
View(IowaIll_lakes)

tally(IowaIll_lakes)
tally(minnesota_lakes)

###There are 16,466 lakes in Iowa and Illinois as opposed to 29,038 in Minnesota

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}

###Outline Iowa
iowa <- states %>%
  filter(name == "Iowa") %>%
  st_transform(2163)

###Clip lakes that occur in Iowa
iowa_lakes <- spatial_lakes[iowa, ] %>%
  mutate(state = "Iowa")


###Create dataframe with Iowa and Minnesota lakes
iowa_min <- rbind(iowa_lakes, minnesota_lakes)

###Plot histogram of lakes size in both states
iowa_min %>%
  ggplot(., aes(lake_area_ha)) +
  geom_histogram(size = 1.5, bins = 25, fill = "blue", color = "black") +
  scale_x_log10() +
  theme_few() +
  theme(text = element_text(size = 20, color = "black")) +
  xlab("Lake Area (ha)") +
  ylab("Count") +
  facet_wrap(~state, dir = 'v') 
  
  
```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}

##Map lakes in Iowa and Illionois and color them by area
IowaIll_lakes %>%
mapview(., zcol = "lake_area_ha")

```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

```{r}
## We could use the National Hydrography data to map lakes and resevoirs and calculuate the lake surface area
```



